---
title: "Practical Machine Learning PeerAssignment"
author: "DivyaAkella"
date: "September 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this project is to predict the manner in which a certain group of people did exercise."classe" variable in the training set has movement category (sitting-down, standing-up, standing, walking, and sitting).TO create a report describing how to built  model, how cross validation is used, the expected out of sample error is. A prediction model is used to predict 20 different test cases.

```{r downloadInputFiles}

setwd("C:/Users/212470996/Desktop/Master_Repo/datasciencecoursera/PracticalMachineLearning")

if(!file.exists("./plmTraining.csv"))
{
  URLTraining <-  "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  download.file(URLTraining,destfile = "plmTraining.csv")
  
}

 plmTraining <- read.csv("./plmTraining.csv", header = TRUE, na.strings = c("NA",""))
 dim(plmTraining)
 
if(!file.exists("./plmTesting.csv"))
{
URLTesting <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(URLTesting,destfile = "plmTesting.csv")

}
plmTesting <- read.csv("./plmTesting.csv", header = TRUE, na.strings = c("NA",""))
dim(plmTesting)

```

## Preprocessing Data

Remove the columns which have 70% of data missing and columns that are not needed

```{r Preprocessing}
library(caret)

plmTraining_Rm_NoDataClmns <- plmTraining[,(colSums(is.na(plmTraining))<0.7*nrow(plmTraining))] 
plmTesting_final <- plmTesting[,(colSums(is.na(plmTesting))<0.7*nrow(plmTesting))]

plmTraining_final <- plmTraining_Rm_NoDataClmns[,-c(1:7)]

  
```

## Get Validation Dataset & Training Dataset

```{r CreateTrainAndValidationDatasets}
library(caret)
inTrain <- createDataPartition(y= plmTraining_final$classe, p= 0.7 , list = FALSE)
Training <- plmTraining_final[inTrain,]
Validation <- plmTraining_final[-inTrain,]

```


## Apply Random Forest method
```{r RamdomForestPredict}

   
  set.seed(123)
modFit_rf <- train(classe ~ ., method = "rf", data = Training,trControl =    trainControl(method = "cv", number = 4, allowParallel = TRUE, verboseIter=TRUE))   
   
   Validation_pred_rf <- predict(modFit_rf,Validation)  
   confusionMatrix(Validation_pred_rf,Validation$classe)

```

 
##Predict Testing Data

```{r PredictTestData}


Testing_Pred <- predict(modFit_rf,plmTesting_final)
Testing_Pred


```
